---
title: 'Part 2: chi-square with lizards''tails'''
author: "Narjes Mathlouthi"
date: "11/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)

```

## Read in data

```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species :

```{r}

lizards %>% 
  count(spp) %>% 
  arrange(-n)
```


## Make a subset 

-USTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST (east tarbush site)
- Tail condition is recorded( W = whole, B = broken)
```{r}
utst <-  lizards %>% 
  filter(spp == "UTST") %>% 
  filter (site %in% c("IBPE", "RABB", "EAST")) %>% 
#unique(utst$site)
  filter(tail %in% c("B", "W"))
```

## Find counts and  proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
# count by site and tail
```


```{r}
tail_counts <- utst %>% 
  janitor :: tabyl(site, tail)
#View(tail_counts)

tail_proportions <- tail_counts %>% 
  # default is count(proportions) percentages per row
  janitor :: adorn_percentages() %>% 
# format by rounding to 2 digits
  janitor :: adorn_pct_formatting(digits = 2) %>% 
  #add counts next to %
  janitor:: adorn_ns()
```


## Convert to a contigency table 

Use `column_to_rowname()`. 

```{r}

# creates a table without storing site name only contains counts
# it is useful for chi square calc
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")


```

## Chi-square

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)

```

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) =  `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).


## END Part 2
